\documentclass[12pt]{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[style=english]{csquotes}
\usepackage[english, spanish, catalan]{babel}
\usepackage{graphicx}
\usepackage{geometry}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
\usepackage{titlesec}
\usepackage{lipsum}
\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz-cd}
\usepackage{subcaption}
\usepackage{stmaryrd}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{shapes,
                tikzmark}
\tikzset{every tikzmarknode/.style={%
        draw=red, semithick, inner sep=2pt}
        }
\usepackage{forest}
\usetikzlibrary{fit,positioning}

\tikzset{
  red arrow/.style={
    midway,red,sloped,fill, minimum height=3cm, single arrow, single arrow head extend=.5cm, single arrow head indent=.25cm,xscale=0.3,yscale=0.15,
    allow upside down
  },
  black arrow/.style 2 args={-stealth, shorten >=#1, shorten <=#2},
  black arrow/.default={1mm}{1mm},
  tree box/.style={draw, rounded corners, inner sep=0.5em},
  node box/.style={white, draw=black, text=black, rectangle, rounded corners},
}
\input{definitions.tex}

%% Bibliography
\usepackage[backend=biber,style=numeric]{biblatex}
\DefineBibliographyStrings{english}{
  bibliography = {References},
}
\addbibresource{references.bib}

%% Cover, abstract and page style
\input{template}

%% Assemble the document
\begin{document}

\setlength{\headheight}{14.49998pt}
\addtolength{\topmargin}{-2.49998pt}

\createcover
\pagenumbering{roman}
\createabstract{english}{The detection of tumoural cells from whole slide images is an essential task in medical diagnosis and research. In this thesis, we propose and analyse a novel approach that combines vision-based models with graph neural networks to improve the accuracy of automated tumoural cell detection. Our proposal leverages the inherent structure and relationships between cells in the tissue. Experimental results on our own curated dataset shows that several different metrics improve by up to $15\%$ compared to just using the vision approach. It has been proved to work with H\&E stained lung tissue and HER2 stained breast tissue. We believe that our proposed method has the potential to improve the accuracy of automated tumoural cell detection, which can lead to accelerated diagnosis and research in the field by reducing the worload of hystopathologists.}{\textbf{Keywords:} Histology, Lung, Breast, Graph Neural Networks, Convolutional Neural Networks, Calibration.}{\textbf{AMS Codes:} 92C50, 92C55, 92C37, 68T10, 68T45, 68R10}
\clearpage
\createabstract{spanish}{La detección de células tumorales en imágenes de portaobjeto completo juega un papel esencial en el diagnóstico médico y es un elemento fundamental de la investigación sobre el cáncer. En esta tesis proponemos y analizamos un enfoque novedoso que combina modelos de visión por ordenador con redes neuronales en grafos para mejorar la precisión de la detección automatizada de células tumorales. Nuestra propuesta aprovecha la estructura inherente y las relaciones entre las células del tejido. Los resultados experimentales obtenidos sobre nuestra propia base de datos muestran que varias métricas mejoran hasta en un 15\% en comparación con solo usar el enfoque de visión. Se ha demostrado que funciona con tejido pulmonar teñido con H\&E y tejido mamario teñido con HER2. Creemos que nuestro método tiene el potencial de mejorar la precisión de los métodos automáticos de detección de células tumorales, lo que puede llevar a acelerar los diagnósticos y la investigación en este ámbito al reducir la carga de trabajo de los histopatólogos.}{\textbf{Palabras clave:} Histología, Pulmón, Mama, Redes Neuronales en Grafos, Redes Neuronales Convolucionales, Calibración.}{\textbf{Códigos AMS:} 92C50, 92C55, 92C37, 68T10, 68T45, 68R10}
\clearpage
\createabstract{catalan}{La detecció de cèl·lules tumorals en imatges de seccions completes és una tasca essencial en el diagnòstic mèdic i la investigació. En aquesta tesi, proposem i analitzem un enfocament innovador que combina models basats en visió amb xarxes neuronals en grafs per millorar la precisió de la detecció automatitzada de cèl·lules tumorals. La nostra proposta aprofita l'estructura inherent i les relacions entre cèl·lules en el teixit. Els resultats experimentals en el nostre propi conjunt de dades curat mostrin que diversos indicadors milloren fins a un 15\% en comparació amb només usar l'enfocament de visió. S'ha demostrat que funciona amb teixit pulmonar tenyit amb H\&E i teixit mamari tenyit amb HER2. Creiem que el nostre mètode proposat té el potencial de millorar la precisió de la detecció automatitzada de cèl·lules tumorals, el que pot portar a uns diagnòstics més ràpids i una investigació accelerada en el camp degut a la reducció en la càrrega de treball dels histopatòlegs.}{\textbf{Paraulas clau:} Histologia, Pulmó, Mama, Xarxes Neuronals en Grafs, Xarxes Neuronals Convolucionals, Calibració.}{\textbf{Codis AMS:} 92C50, 92C55, 92C37, 68T10, 68T45, 68R10}
\tableofcontents
\clearpage
\pagenumbering{arabic}

\input{content/Introduction}
\input{content/Problem Formulation}
\input{content/Problem Solving}
\input{content/Results}
\input{content/Conclusions}
\input{content/Future work}
\printbibliography[heading=bibintoc]
\input{content/Appendix}

\end{document}
